### 2015-01-14 fundamentals/impl.tex
comma
todo cleanup

In large simulations many material objects may change frequently.
Material decay can also contribute significantly to such changes.  In order to
help avoid unnecessary runtime performance and database space impacts,
compositions in \Cyclus have some special features.
Immutability of compositions allows multiple material objects to all hold
references to the same composition object safely.  When operations are
performed on a material that do not alter the composition, any new resulting
materials hold a reference to the same, original composition object.
Although the decay operation generates a new, separate composition (preserving
immutability), a reference to the new composition is placed in a cache shared
by the new and original compositions. The next time the original
composition is decayed, this cache will be checked for a prior calculation
with the same cumulative decay time.  If one exists, the cached composition is
used.  This cache is shared by all compositions that transitively come from
the same origin.  Composition immutability in concert with decay history
caching help eliminate many redundant decay calculations in addition to
reducing the total number of composition objects.  Since each composition
object is only recorded in the database once, significant space savings also
occurs. \TODO{do we want a diagram helping show this composition caching?}

@@

The \Cyclus toolkit provides two useful interfaces for queries referring to physical paramaters of \Class{Material}
objects. First, the \Class{MatQuery} module
provides a basic querying \gls{API}, including the atom and mass fractions of
nuclides, the number of moles of a nuclide in a material, and also the amount of
aggregate nuclides, i.e., a \Class{Composition}, in a material. The
\Class{Enrichment} module provides an \gls{API} for determining enrichment-related
parameters of a material, including the separative work units (SWU) and natural
uranium required to enrich a material provided knowledge of feed, product, and tails
assays.

@@

In addition to those that already exist, new tools and computational patterns will 
emerge from the archetypes that are developed within the community of 
developers. As these tools gain adoption between projects and demonstrate their 
utility to the developer community, they will be considered for screening and 
adoption into the kernel as toolkit extensions. Likely extensions include:

>>>>

In large simulations, many material objects may change frequently.
Material decay can also contribute significantly to such changes.  In order to
help avoid unnecessary runtime performance and database space impacts,
compositions in \Cyclus have some special features.
Immutability of compositions allows multiple material objects to all hold
references to the same composition object safely.  When operations are
performed on a material that do not alter the composition, any new resulting
materials hold a reference to the same, original composition object.
Although the decay operation generates a new, separate composition (preserving
immutability), a reference to the new composition is placed in a cache shared
by the new and original compositions. The next time the original
composition is decayed, this cache will be checked for a prior calculation
with the same cumulative decay time.  If one exists, the cached composition is
used.  This cache is shared by all compositions that transitively come from
the same origin.  Composition immutability in concert with decay history
caching help eliminate many redundant decay calculations in addition to
reducing the total number of composition objects.  Since each composition
object is only recorded in the database once, significant space savings also
occurs. \TODO{do we want a diagram helping show this composition caching?}

@@

The \Cyclus toolkit provides two useful interfaces for querying physical paramaters of \Class{Material}
objects. First, the \Class{MatQuery} module
provides a basic querying \gls{API}, including the atom and mass fractions of
nuclides, the number of moles of a nuclide in a material, and also the amount of
aggregate nuclides, i.e., a \Class{Composition}, in a material. The
\Class{Enrichment} module provides an \gls{API} for determining enrichment-related
parameters of a material, including the separative work units (SWU) and natural
uranium required to enrich a material provided knowledge of feed, product, and tails
assays.

@@

In addition to those that already exist, new tools and computational patterns will 
emerge from the archetypes developed within the community of 
developers. As these tools gain adoption between projects and demonstrate their 
utility to the developer community, they will be considered for screening and 
adoption into the kernel as toolkit extensions. Likely extensions include:
